### 2022-04-05 outputs/paper/paper.Rmd
Updated Abstract and Introduction sections of the paper.

---
title: "(TEMP) Online Video Game Monetization Schemes and Its Relationship to the Consumers"
author: 
  - Tian Yi Zhang
thanks: "Code and data are available at: https://github.com/zhan7818/bdo_monetization"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Advertisement and monetization schemes are essential to obtaining a stable revenue stream and ensuring the continued surival of an online video game. We obtain data through a survey distributed in several online communities for the massively multiplayer online game known as Black Desert Online and analyze it throuhg several graphs, tables, and a model in this paper. (TODO) Sentence about findings. (TODO) Sentence about implications. (TODO) Keywords:"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Introduction

Just a decade or two ago, singleplayer-driven content was the predominant
focus in the video game industry. This was mostly a product of the time, as the
infrastructures needed to support online-based video games were yet to be
fully developed. More importantly, what this also meant was a lack of 
post-launch updates being delivered for a game, as developers also lacked means
to deliver constant updates effectively into the hands of the consumers. As a
result, most video games received little to no post-launch updates aside from
one or two major expansions to their games that contained sufficient content
to be put on the shelves as add-ons to the original purchase.
Despite this, consumers back then were generally content with the amount of 
content they received for their purchases.

As the years went by and networking structures matured, however, more and more
video game developers started focusing more on the online aspects of gaming.
The number of multiplayer-focused video games in the market started to steadily
increase. Now, in an era of live service online video games, developers are 
pressured to continuously produce and deliver new and exciting content on a 
consistent schedule in order to compete with other products. This need to
create more content for a game means that development resources are constantly
required, and online focused games can no longer afford to charge their
consumers a simple upfront fee; live service video games must find an effective
and profitable monetization scheme in order to fund their continued development.

The most prominent examples of live service online video games are the massively
multiplayer online role-playing games (MMORPG), a genre of video games known
specifically for its focus on multiplayer. Due to their focus on 
progression-based gameplay, developers of MMORPGs must constantly create new
and refreshing content to add to their games. To subsidize the development
costs that go along with it, developers have created a variety of monetization
models to ensure a steady revenue stream. For example, World of Warcraft, one
of the biggest MMORPGs, relies primarily on a monthly subscription from its
players. Others, such as Lost Ark, adopt a mix of strategies with 
a monthly subscription that grants
massive boosts in-game and an in-game cash shop that allows players to
purchases items for a variety of purposes using real world currencies.
If done properly, an analysis
on the monetization schemes of MMORPGs could have meaningful
implications for future MMORPG developers on how they should monetize their
video games to maximize revenue.

This paper primarily focuses on the MMORPG known as Black Desert Online,
as it is MMORPG that utilizes a of mix of upfront cost, subscription and in-game
cash shop. Pearl Abyss, the developers behind Black Desert Online, 
never published any official statistics that touch upon the revenue of
Black Desert Online. As such, the following analysis will rely on data 
collected by the author of this paper.

The remainder of this paper is organized as follows: Section \@ref(data) covers
the survey design, sampling and preliminary analyses of the dataset used in this
paper. Section \@ref(model) details the model used to find possible correlations
between variables in the dataset, the results and implications of which will 
be covered in Section \@ref(results). Any limitations or possible extensions 
of this paper will be discussed in Section \@ref(discussion).

TODO: add in some references that support claims about general transition from 
singleplayer to multiplayer in video game industry.

# Data

TODO

# Model

TODO

# Results

TODO

# Discussion

TODO

\newpage

\appendix

# Appendix {-}

\newpage

# References